---
title: "Bootstrap Geweke Diagnostic: An Introduction"
author: "Jay"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bootstrap Geweke Diagnostic: An Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bootGeweke)
```

# Introduction

This vignette demonstrates how to use the `bootstrap_geweke()` function from the **bootGeweke** package to perform a Geweke diagnostic with bootstrap sampling.

The Geweke diagnostic is a convergence diagnostic for Markov Chain Monte Carlo (MCMC) chains. This package introduces a bootstrap approach to calculate Geweke Z-scores and confidence intervals.

## Installing the Package

You can install the development version of the package using:

```{r example_geweke_diagnostic, echo = TRUE}
library(bootGeweke)
library(coda)

# Generate a synthetic MCMC chain
set.seed(123)
example_chain <- coda::as.mcmc.list(coda::mcmc(rnorm(1000)))

# Perform the bootstrap Geweke diagnostic
result <- bootstrap_geweke(example_chain, B = 500, n = 500, confidence_level = 0.95, frac1 = 0.1, frac2 = 0.5)

# Print the results
print(result)
```

## Plotting the Z-scores

The `plot()` method for `bootGeweke` objects allow you to visualize the distribution of Z-scores:       

```{r}
plot(result)
```

## Summary the Z-scores statistic       

The `summary()` method for `bootGeweke` objects allow you to summaries the statistics of Z-scores distribution: `Mean Z-score`, `Standard Deviation`, `Confidence Intervals`.

```{r}
summary(result)
```

## Handling Input Erros

The function includes several input checks to ensure valid inputs. For example:

```{r}
# Invalid input type
tryCatch({
  bootstrap_geweke("invalid_input")
}, error = function(e) {
  print(e)
})
```

## Conclusion       

The `bootstrap_geweke()` function provides an effective way to perform a bootstrap-based Geweke diagnostic on MCMC chains. It allows users to compute confidence intervals for Z-scores, improving the interpretation of convergence diagnostics in MCMC simulations.       
























